### 2018-02-07 submission/manuscript.Rmd
update genera based on OR

***Mouth-Associated Genera are Associated with an Increased OR of Tumor:***
Next, we asked if being higher than the median relative abundance, for any specific genera, resulted in an altered OR for adenoma or carcinoma, in stool and colon tissue, due to our previous observations of small increases in OR using OTU richness and Shannon diversity. To investigate this we analyzed all common genera across each study, in colon tissue or stool, and assessed whether a relative abundance higher than the median results in an increase or decrease in RR. Mouth-associated genera were commonly found in the top 5 genera associated with an increased OR of having an adenoma (*Pyramidobacter* [Figure 3A] and *Rothia* [Figure 3C]) and carcinoma (*Fusobacterium*, *Parvimonas*, *Porphyromonas*, and *Peptostreptococcus* [Figure 3B] and *Fusobacterium* [Figure 3D]) for both stool and colon tissue samples. Conversely, genera commonly associated with a normal gastrointestinal tract were correlated with a decreased OR for both adenoma and carcinoma for both stool and colon tissue samples [Figure 3]. Even though mouth-associated genera were identified across disease stage, there was little direct overlap of the top 5 increased or decreased OR genera between both stages and sample site.

When observing ORs with a P-value less than 0.05 there was almost no overlap between genera from stool or colon tissue and when they were similar the RR was in opposite directions (e.g. *Lactococcus*) [Table S8 & S9]. Many of the genera that had ORs with a P-value under 0.05 for colon tissue are also highly prevalent in contamination, specifically, *Novosphingobium*, *Selemonas*, and *Achromobacter* [Figure 3 & Table S8-S9]. For carcinoma stage of disease, certain mouth-associated genera (*Fusobacterium*, *Parvimonas*, *Peptostreptococcus*) had a high OR for both colon tissue and stool samples [Table S10 & S11]. Finally, these data suggest that the most significantly increased RR genera for tissue was *Camplyobacter* while in stool it was *Peptostreptococcus* [Table S10 & S11].

@@

**Figure 3: Top 5 Genera that Decrease and Increase Odds Ratio for Lesion.** A) Adenoma odds ratio in stool. B) Carcinoma odds ratio in stool. C) Adenoma odds ratio in tissue. D) Carcinoma odds ratio in tissue. For all panels the odds ratio was also compared to whether one, two, three, or four of the CRC-associated genera were present.

>>>>

***Mouth-Associated Genera are Associated with an Increased OR of Tumor:***
Next, we asked if being higher than the median relative abundance, for any specific genera, resulted in an altered OR for adenoma or carcinoma, in stool and colon tissue, due to our previous observations of small increases in OR using OTU richness and Shannon diversity. To investigate this we analyzed all common genera across each study, in colon tissue or stool, and assessed whether a relative abundance higher than the median results in an increase or decrease in RR. Mouth-associated genera were commonly found in the top 5 genera associated with an increased OR of having an adenoma (*Porphyromonas* [Figure 3A] and *Rothia* [Figure 3C]) and carcinoma (*Fusobacterium*, *Parvimonas*, *Porphyromonas*, and *Peptostreptococcus* [Figure 3B] and *Fusobacterium* [Figure 3D]) for both stool and colon tissue samples. Conversely, genera commonly associated with a normal gastrointestinal tract were correlated with a decreased OR for both adenoma and carcinoma for both stool and colon tissue samples [Figure 3]. Even though mouth-associated genera were identified across disease stage, there was little direct overlap of the top 5 increased or decreased OR genera between both stages and sample site.

When observing ORs with a P-value less than 0.05 there was almost no overlap between genera from stool or colon tissue and when they were similar the RR was in opposite directions (e.g. *Lactococcus*) [Table S8 & S9]. Many of the genera that had ORs with a P-value under 0.05 for colon tissue are also highly prevalent in contamination, specifically, *Novosphingobium*, *Pseudomonas*, and *Achromobacter* [Figure 3 & Table S8-S9]. For carcinoma stage of disease, certain mouth-associated genera (*Fusobacterium*, *Parvimonas*) had an increased OR for both colon tissue and stool samples [Table S10 & S11]. Finally, these data suggest that the most significantly increased OR genera for tissue was *Lepttrichia* while in stool it was *Peptostreptococcus* [Table S10 & S11].

@@

**Figure 3: Top 5 Genera that Decrease and Increase Odds Ratio for Lesion.** A) Adenoma odds ratio in stool. B) Carcinoma odds ratio in stool. C) Adenoma odds ratio in tissue. D) Carcinoma odds ratio in tissue. For all panels the odds ratio was also compared to whether one, two, three, or four of the CRC-associated genera were present. Points represented as only half on the graph have an OR of infinity in the positive or negative direction.

### 2018-02-07 submission/manuscript.tex
update genera based on OR

\textbf{\emph{Mouth-Associated Genera are Associated with an Increased
OR of Tumor:}} Next, we asked if being higher than the median relative
abundance, for any specific genera, resulted in an altered OR for
adenoma or carcinoma, in stool and colon tissue, due to our previous
observations of small increases in OR using OTU richness and Shannon
diversity. To investigate this we analyzed all common genera across each
study, in colon tissue or stool, and assessed whether a relative
abundance higher than the median results in an increase or decrease in
RR. Mouth-associated genera were commonly found in the top 5 genera
associated with an increased OR of having an adenoma
(\emph{Pyramidobacter} {[}Figure 3A{]} and \emph{Rothia} {[}Figure
3C{]}) and carcinoma (\emph{Fusobacterium}, \emph{Parvimonas},
\emph{Porphyromonas}, and \emph{Peptostreptococcus} {[}Figure 3B{]} and
\emph{Fusobacterium} {[}Figure 3D{]}) for both stool and colon tissue
samples. Conversely, genera commonly associated with a normal
gastrointestinal tract were correlated with a decreased OR for both
adenoma and carcinoma for both stool and colon tissue samples {[}Figure
3{]}. Even though mouth-associated genera were identified across disease
stage, there was little direct overlap of the top 5 increased or
decreased OR genera between both stages and sample site.

When observing ORs with a P-value less than 0.05 there was almost no
overlap between genera from stool or colon tissue and when they were
similar the RR was in opposite directions (e.g. \emph{Lactococcus})
{[}Table S8 \& S9{]}. Many of the genera that had ORs with a P-value
under 0.05 for colon tissue are also highly prevalent in contamination,
specifically, \emph{Novosphingobium}, \emph{Selemonas}, and
\emph{Achromobacter} {[}Figure 3 \& Table S8-S9{]}. For carcinoma stage
of disease, certain mouth-associated genera (\emph{Fusobacterium},
\emph{Parvimonas}, \emph{Peptostreptococcus}) had a high OR for both
colon tissue and stool samples {[}Table S10 \& S11{]}. Finally, these
data suggest that the most significantly increased RR genera for tissue
was \emph{Camplyobacter} while in stool it was \emph{Peptostreptococcus}
{[}Table S10 \& S11{]}.

@@

\textbf{Figure 3: Top 5 Genera that Decrease and Increase Odds Ratio for
Lesion.} A) Adenoma odds ratio in stool. B) Carcinoma odds ratio in
stool. C) Adenoma odds ratio in tissue. D) Carcinoma odds ratio in
tissue. For all panels the odds ratio was also compared to whether one,
two, three, or four of the CRC-associated genera were present.

>>>>

\textbf{\emph{Mouth-Associated Genera are Associated with an Increased
OR of Tumor:}} Next, we asked if being higher than the median relative
abundance, for any specific genera, resulted in an altered OR for
adenoma or carcinoma, in stool and colon tissue, due to our previous
observations of small increases in OR using OTU richness and Shannon
diversity. To investigate this we analyzed all common genera across each
study, in colon tissue or stool, and assessed whether a relative
abundance higher than the median results in an increase or decrease in
RR. Mouth-associated genera were commonly found in the top 5 genera
associated with an increased OR of having an adenoma
(\emph{Porphyromonas} {[}Figure 3A{]} and \emph{Rothia} {[}Figure 3C{]})
and carcinoma (\emph{Fusobacterium}, \emph{Parvimonas},
\emph{Porphyromonas}, and \emph{Peptostreptococcus} {[}Figure 3B{]} and
\emph{Fusobacterium} {[}Figure 3D{]}) for both stool and colon tissue
samples. Conversely, genera commonly associated with a normal
gastrointestinal tract were correlated with a decreased OR for both
adenoma and carcinoma for both stool and colon tissue samples {[}Figure
3{]}. Even though mouth-associated genera were identified across disease
stage, there was little direct overlap of the top 5 increased or
decreased OR genera between both stages and sample site.

When observing ORs with a P-value less than 0.05 there was almost no
overlap between genera from stool or colon tissue and when they were
similar the RR was in opposite directions (e.g. \emph{Lactococcus})
{[}Table S8 \& S9{]}. Many of the genera that had ORs with a P-value
under 0.05 for colon tissue are also highly prevalent in contamination,
specifically, \emph{Novosphingobium}, \emph{Pseudomonas}, and
\emph{Achromobacter} {[}Figure 3 \& Table S8-S9{]}. For carcinoma stage
of disease, certain mouth-associated genera (\emph{Fusobacterium},
\emph{Parvimonas}) had an increased OR for both colon tissue and stool
samples {[}Table S10 \& S11{]}. Finally, these data suggest that the
most significantly increased OR genera for tissue was \emph{Lepttrichia}
while in stool it was \emph{Peptostreptococcus} {[}Table S10 \& S11{]}.

@@

\textbf{Figure 3: Top 5 Genera that Decrease and Increase Odds Ratio for
Lesion.} A) Adenoma odds ratio in stool. B) Carcinoma odds ratio in
stool. C) Adenoma odds ratio in tissue. D) Carcinoma odds ratio in
tissue. For all panels the odds ratio was also compared to whether one,
two, three, or four of the CRC-associated genera were present. Points
represented as only half on the graph have an OR of infinity in the
positive or negative direction.
